---
title: "Uber Data Analysis"
output: html_document
---

##Importing Packages [ggplot2, ggthemes, lubridate, dplyr, tidyr, DT, scales]
```{r}

library(ggplot2)
library(ggthemes)
library(lubridate)
library(dplyr)
library(tidyr)
library(DT)
library(scales)
```


##Creating vector of colors for the plots
```{r}

colors = c("#2f2d2d", "#eaeaea", "#05a399", "#7676a7", "#f5e840", "#ab47bc", "#388e3c")
colors
```
   
##Current Working Directory
```{r}
getwd()
```

##Reading the data for each month separately 
```{r}
apr <- read.csv("C:/Users/py96230n/Desktop/Uberdata/uber-raw-data-apr14.csv")
may <- read.csv("C:/Users/py96230n/Desktop/Uberdata/uber-raw-data-may14.csv")
june <- read.csv("C:/Users/py96230n/Desktop/Uberdata/uber-raw-data-jun14.csv")
july <- read.csv("C:/Users/py96230n/Desktop/Uberdata/uber-raw-data-jul14.csv")
aug <- read.csv("C:/Users/py96230n/Desktop/Uberdata/uber-raw-data-aug14.csv")
sept <- read.csv("C:/Users/py96230n/Desktop/Uberdata/uber-raw-data-sep14.csv")

```


##Combine the data together 
#rbind multiple group of rows
```{r}
data <- rbind(apr, may, june, july, aug, sept)
cat("The dimensions of the data are:", dim(data))
```
```{r}
head(data)
```
##Readable format for the DateTime
```{r}
data$Date.Time <- as.POSIXct(data$Date.Time, format="%m/%d/%Y %H:%M:%S")
data$Time <- format(as.POSIXct(data$Date.Time, format = "%m/%d/%Y %H:%M:%S"), format="%H:%M:%S")
data$Date.Time <- ymd_hms(data$Date.Time)
head(data)

```

#Create individual columns for month, day and year
```{r}
data$day <- factor(day(data$Date.Time))
data$month <- factor(month(data$Date.Time, label=TRUE))
data$year <- factor(year(data$Date.Time))
data$dayofweek <- factor(wday(data$Date.Time, label=TRUE))
head(data)
```
#Adding Time variables
```{r}

data$second = factor(second(hms(data$Time)))
data$minute = factor(minute(hms(data$Time)))
data$hour = factor(hour(hms(data$Time)))
head(data)
```

##Data Visualisation

#Plotting the trips by hours in a day

```{r}
hourly_data <- data %>% 
  group_by(hour) %>% 
  dplyr::summarize(Total = n())
hourly_data
```

```{r}
datatable(hourly_data)
```

#Plot the data by hour
```{r}
ggplot(hourly_data, aes(hour, Total)) + 
  geom_bar(stat="identity", 
           fill="purple", 
           color="yellow") + 
  ggtitle("Trips Every Hour") + 
  theme(legend.position = "none", 
        plot.title = element_text(hjust = 0.5), 
        plot.subtitle = element_text(hjust = 0.5)) + 
  scale_y_continuous(labels=comma)


```
##Plotting trips by hour and month
#Aggregate the data by month and hour
```{r}
month_hour_data <- data %>% group_by(month, hour) %>%  dplyr::summarize(Total = n())
```
```{r}
ggplot(month_hour_data, aes(hour, Total, fill=month)) + 
  geom_bar(stat = "identity") + 
  ggtitle("Trips by Hour and Month") + 
  scale_y_continuous(labels = comma)
```
#Plotting data by trips during every day of the month
```{r}

day_data <- data %>% group_by(day) %>% dplyr::summarize(Trips = n())
day_data
```
#Data for the day
```{r}
ggplot(day_data, aes(day, Trips)) + 
  geom_bar(stat = "identity", fill = "green") +
  ggtitle("Trips by day of the month") + 
  theme(legend.position = "none") + 
  scale_y_continuous(labels = comma)

```


##Collect data by day of the week and month
```{r}
day_month_data <- data %>% group_by(dayofweek, month) %>% dplyr::summarize(Trips = n())
day_month_data
```

#Plot the Collect data by day of the week and month
```{r}
ggplot(day_month_data, aes(dayofweek, Trips, fill = month)) + 
  geom_bar(stat = "identity", aes(fill = month), position = "dodge") + 
  ggtitle("Trias by Day and Month") + 
  scale_y_continuous(labels = comma) + 
  scale_fill_manual(values = colors)
```
#Number of Trips place during months in a year
```{r}
month_data <- data %>% group_by(month) %>% dplyr::summarize(Total = n())
month_data
```

#Plot the Number of Trips place during months in a year
```{r}
ggplot(month_data, aes(month, Total, fill = month)) + 
  geom_bar(stat = "Identity") + 
  ggtitle("Trips in a month") + 
  theme(legend.position = "none") + 
  scale_y_continuous(labels = comma) + 
  scale_fill_manual(values = colors)

```

##Heatmap visualization of day, hour and month
#Heatmap by Hour and Day
```{r}

day_hour_data <- data %>% group_by(day, hour) %>% dplyr::summarize(Total = n())
datatable(day_hour_data)
```

##Heatmap by Hour and Day
```{r}
ggplot(day_hour_data, aes(day, hour, fill = Total)) + 
  geom_tile(color = "white") + 
  ggtitle("Heat Map by Hour and Day")

```

##Heatmap by Day and Month
```{r}

month_day_data <- data %>% group_by(month, day) %>% dplyr::summarize(Trips = n())
month_day_data
```

##Heatmap by Collecting data by month and day
```{r}

ggplot(month_day_data, aes(day, month, fill = Trips)) + 
  geom_tile(color = "white") + 
  ggtitle("Heat Map by Month and Day")

```

##Heatmap by day of the week and month
```{r}
ggplot(day_month_data, aes(dayofweek, month, fill = Trips)) + 
  geom_tile(color = "white") + 
  ggtitle("Heat Map by Month and Day")

```

#Creating a map visualization of rides in NYC
#Set Map Constants
```{r}

min_lat <- 40 
max_lat <- 40.91
min_long <- -74.15
max_long <- -73.7004

ggplot(data, aes(x=Lon, y=Lat)) +
  geom_point(size=1, color = "blue") +
  scale_x_continuous(limits=c(min_long, max_long)) +
  scale_y_continuous(limits=c(min_lat, max_lat)) +
  theme_map() +
   ggtitle("NYC MAP ON UBER RIDES DURING 2014 (APR-SEP)")
```

```{r}

ggplot(data, aes(x=Lon, y=Lat, color = Base)) +
  geom_point(size=1) +
  scale_x_continuous(limits=c(min_long, max_long)) +
  scale_y_continuous(limits=c(min_lat, max_lat)) +
  theme_map() +
  ggtitle("NYC MAP ON UBER RIDES DURING 2014 (APR-SEP) by BASE")
```

